Optimal updating magnitude in adaptive flat-distribution-sampling simulations

### 2017-09-14 doc/invt.tex
improve reweight.c and web

\item \label{step:Gamma}
Estimate $\Gamma_k$'s from Eq. \eqref{eq:varXt}
(cf. Appendix \ref{sec:Gamma_measure})
as well as the initial bias $\epsilon_k$.\footnote{The
  value of $\tilde u_k$ can be estimated
  from the histogram-corrected bias potential
  by Eq. \eqref{eq:uhatav},
  although it often leads to overestimated
  $\epsilon_k$'s for large-$k$ modes.
}

@@

The schedule, $\alpha(t)$,
also defines a variable transformation from
the mass distribution, $m(\bar q)$.
%much like the Legendre transform.
%
If we define
$\bar q^* \equiv \bar q/q(T)$
and $t^* \equiv t/T$
such that both are limited to the interval $[0, 1]$,
as well as
\begin{align*}
  m^*       &\equiv q(T) \, m(\bar q) = -\frac{ dt^* } { d\bar q^* },
  \\
  \alpha^*  &\equiv \frac{T}{q(T)} \, \alpha(t) = -\frac{ d\bar q^* } { dt^* },
\end{align*}
then it is clear that
both $m^*(\bar q^*)$ and $\alpha^*(t^*)$
can describe the differential relation
between $\bar q^*$ and $t^*$,
but with the natural variables
being $\bar q^*$ and $t^*$, respectively.
%
Further, the transformation is an involution:
if we apply the same transform on $\alpha^*(t^*)$
we would recover $m^*(\bar q^*)$,
as shown in Fig. \ref{fig:massq}(b).
%
A few examples are shown in Table \ref{tab:m_and_a}.
%
For example, in the single-bin updating scheme,
the mass distribution decays exponentially,
and we would have the inverse-time schedule.
%
On the other hand,
the eigenvalues of the Gaussian updating scheme
can differ greatly in magnitude
[cf. Eq. \eqref{eq:lambda_Gaussian}].
%
Then, at a large $\bar q$,
the sum in Eq. \eqref{eq:mass_func} will be dominated by
the largest term, where $\lambda_k \approx 1/\bar q$,
and $m(\bar q) \sim 1/\bar q$,
and we would expect a roughly exponentially-decaying schedule.

@@

Consider an FDS simulation
under a nearly exact bias potential.
%
By Eq. \eqref{eq:sh_ave},
the fluctuation of the instantaneous histogram
is then dominated by the random part,
%
\begin{equation}
  f_{*i}(t) \approx \xi_{*i}(t)
  ,
  \label{eq:f_xi}
\end{equation}
and
%the same holds for the Fourier transform,
$\tilde f_k(t) \approx \tilde \xi_k(t).$
%
If we define the time average,
%
\begin{align*}
  \tilde F_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde f_k(t)
  =\mathcal F\left[ \frac{ H_i(T) } { \rho_i } \right]_k
  ,
\end{align*}
%
for the histogram fluctuation of mode $k$,
then the total fluctuation is
%
\begin{align}
  F^2(T)
  &\equiv
  \sum_{k=1}^{n-1} \tilde F_k^2(T)
  =
  \sum_{i=1}^{n} \rho_i \, F_{*i}^2(T)
  =
  \sum_{i=1}^n
  \frac{ [ H_i(T) -\rho_i]^2 }{\rho_i}
  .
  \notag
  %\label{eq:hfluc_def}
\end{align}
%
From Eq. \eqref{eq:f_xi}, we have
%$$\bigl\langle \tilde F_k^2(T) \bigr\rangle
%=
%\sum_{t=-\infty}^\infty
%\frac{
%\bigl\langle
%  \tilde f_k(t) \, \tilde f_k(0)
%\bigr\rangle } { T }
%\approx
%\sum_{t=-\infty}^\infty
%\frac{
%\bigl\langle
%  \tilde \xi_k(t) \, \tilde \xi_k(0)
%\bigr\rangle } { T }
%=
%\frac{ \Gamma_k } { T },$$
%
\begin{align}
  \bigl\langle \tilde F_k^2(T) \bigr\rangle
  &=
  \frac{1}{T}
  \sum_{t=-\infty}^\infty
  \bigl\langle
    \tilde f_k(t) \, \tilde f_k(0)
  \bigr\rangle
  \notag \\
  &\approx
  \frac{1}{T}
  \sum_{t=-\infty}^\infty
  \bigl\langle
    \tilde \xi_k(t) \, \tilde \xi_k(0)
  \bigr\rangle
  =
  \frac{ \Gamma_k } { T }
  ,
  \notag
  %\label{eq:Fk2_Gammak}
\end{align}
%
and the total histogram fluctuation is
\begin{equation}
  \bigl\langle F^2(T) \bigr\rangle
  =
  \sum_{k=1}^{n-1}
  \bigl\langle \tilde F_k^2(T) \bigr\rangle
  =
  \sum_{k=1}^{n-1}
  \frac{ \Gamma_k } { T }
  =
  \frac{ \Gamma } { T }
  ,
  \label{eq:F2sum}
\end{equation}
where we have used Eq. \eqref{eq:Gammak_sum}
in the last step.
%
We may relate the histogram fluctuation
to the minimal error under the single-bin updating scheme,
Eq. \eqref{eq:Emin_sbin}, as
  $\Err(T) \approx \bigl\langle F^2(T) \bigr\rangle$.
%we may rephrase the optimality of the inverse-time schedule
%as the inequality,
%\begin{equation}
%  \Err(T)
%  \approx
%  \bigl\langle F^2(T) \bigr\rangle
%  .
%  \notag
%  %\label{eq:fluc}
%\end{equation}
%and the histogram fluctuation
%serves as a good estimate of the final error.
%

>>>>

\item \label{step:Gamma}
Estimate $\Gamma_k$'s
(cf. Appendix \ref{sec:Gamma_measure})
and the initial bias $\epsilon_k$.\footnote{The
  value of $\tilde u_k$ can be estimated
  from the histogram-corrected bias potential
  by Eq. \eqref{eq:uhatav},
  although it often leads to overestimated
  $\epsilon_k$'s for large-$k$ modes.
}

@@

The schedule, $\alpha(t)$,
also defines a variable transformation from
the mass distribution, $m(\bar q)$.
%much like the Legendre transform.
%
If we define
$\bar q^* \equiv \bar q/q(T)$
and $t^* \equiv t/T$
such that both are limited to the interval $[0, 1]$,
as well as
\begin{align*}
  m^*       &\equiv q(T) \, m(\bar q) = -\frac{ dt^* } { d\bar q^* },
  \\
  \alpha^*  &\equiv \frac{T}{q(T)} \, \alpha(t) = -\frac{ d\bar q^* } { dt^* },
\end{align*}
then it is clear that
both $m^*(\bar q^*)$ and $\alpha^*(t^*)$
can describe the differential relation
between $\bar q^*$ and $t^*$,
but with the natural variables
being $\bar q^*$ and $t^*$, respectively.
%
Further, the transformation is an involution:
if we apply the same transform on $\alpha^*(t^*)$
we would recover $m^*(\bar q^*)$,
as shown in Fig. \ref{fig:massq}(b).
%
A few examples are shown in Table \ref{tab:m_and_a}.
%
For example, in the single-bin updating scheme,
the mass distribution decays exponentially,
and we would have the inverse-time schedule.
%
On the other hand,
the eigenvalues of the Gaussian updating scheme
can differ greatly in magnitude
[cf. Eq. \eqref{eq:lambda_Gaussian}].
%
Then, at a large $\bar q$,
the sum in Eq. \eqref{eq:mass_func} will be dominated by
the largest term, where $\lambda_k \approx 1/\bar q$,
and $m(\bar q) \propto M(\bar q) \propto 1/\bar q$,
and we would expect a roughly exponentially-decaying schedule.

@@

Consider an FDS simulation
under a nearly exact bias potential.
%
By Eq. \eqref{eq:sh_ave},
the fluctuation of the instantaneous histogram
is then dominated by the random part,
%
\begin{equation}
  f_{*i}(t) \approx \xi_{*i}(t)
  ,
  \label{eq:f_xi}
\end{equation}
and
%the same holds for the Fourier transform,
$\tilde f_k(t) \approx \tilde \xi_k(t).$
%
If we define the time average,
%
\begin{align*}
  \tilde F_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde f_k(t)
  =\mathcal F\left[ \frac{ H_i(T) } { \rho_i } \right]_k
  ,
\end{align*}
%
for the histogram fluctuation of mode $k$,
then the total fluctuation is
%
\begin{align}
  F^2(T)
  &\equiv
  \sum_{k=1}^{n-1} \tilde F_k^2(T)
  =
  \sum_{i=1}^{n} \rho_i \, F_{*i}^2(T)
  =
  \sum_{i=1}^n
  \frac{ [ H_i(T) -\rho_i]^2 }{\rho_i}
  ,
  \notag
  %\label{eq:hfluc_def}
\end{align}
%
where we have used Eq. \eqref{eq:parseval} in the last step.
%
From Eq. \eqref{eq:f_xi}, we have
%$$\bigl\langle \tilde F_k^2(T) \bigr\rangle
%=
%\sum_{t=-\infty}^\infty
%\frac{
%\bigl\langle
%  \tilde f_k(t) \, \tilde f_k(0)
%\bigr\rangle } { T }
%\approx
%\sum_{t=-\infty}^\infty
%\frac{
%\bigl\langle
%  \tilde \xi_k(t) \, \tilde \xi_k(0)
%\bigr\rangle } { T }
%=
%\frac{ \Gamma_k } { T },$$
%
\begin{align}
  \bigl\langle \tilde F_k^2(T) \bigr\rangle
  &=
  \frac{1}{T}
  \sum_{t=-\infty}^\infty
  \bigl\langle
    \tilde f_k(t) \, \tilde f_k(0)
  \bigr\rangle
  \notag \\
  &\approx
  \frac{1}{T}
  \sum_{t=-\infty}^\infty
  \bigl\langle
    \tilde \xi_k(t) \, \tilde \xi_k(0)
  \bigr\rangle
  =
  \frac{ \Gamma_k } { T }
  ,
  \notag
  %\label{eq:Fk2_Gammak}
\end{align}
%
and the total histogram fluctuation is
\begin{equation}
  \bigl\langle F^2(T) \bigr\rangle
  =
  \sum_{k=1}^{n-1}
  \bigl\langle \tilde F_k^2(T) \bigr\rangle
  =
  \sum_{k=1}^{n-1}
  \frac{ \Gamma_k } { T }
  =
  \frac{ \Gamma } { T }
  ,
  \label{eq:F2sum}
\end{equation}
where we have used Eq. \eqref{eq:Gammak_sum}
in the last step.
%
We may relate the histogram fluctuation
to the minimal error under the single-bin updating scheme,
Eq. \eqref{eq:Emin_sbin}, as
  $\Err(T) \approx \bigl\langle F^2(T) \bigr\rangle$.
%we may rephrase the optimality of the inverse-time schedule
%as the inequality,
%\begin{equation}
%  \Err(T)
%  \approx
%  \bigl\langle F^2(T) \bigr\rangle
%  .
%  \notag
%  %\label{eq:fluc}
%\end{equation}
%and the histogram fluctuation
%serves as a good estimate of the final error.
%
